Concurrent, Parallel Garbage Collection in Linear Time

### 2016-06-02 finaletd/prelim.tex
prelim changes

The heuristic idea mentioned above prevents the creation of strong cycles. The other technique required by the Brownbridge algorithm is the ability to change the strong and weak edges whenever required by the nodes. To solve this issue, Brownbridge used two boolean values, one for each edge in the source and one for each node. For an edge to be strong, both the boolean values have to be same. So node that wants to change the type of the edge, a node changes the corresponding boolean value. A node will change all the incoming edge type when it changes the boolean value of it. A node by changing the boolean value for outgoing edge changes the type of only the edge whose boolean value is changed. The idea of toggling between the edge type is an important piece of the Brownbridge's algorithm.

\section{Brownbridge Garbage Collector}
Brownbridge algorithm works for a system where adversary stops for garbage collection process. Each node contains two counters each corresponding to the type of incoming edge. When the graph mutations are serialized, any addition / deletion of edge increments / decrements the corresponding counters in the target node. An addition event never makes a node garbage. A deletion event might make a node garbage. If the edge deleted is weak, no further actions are taken except decrementing the counter. A last strong incoming edge deletion to a node performs additional work. When the last strong incoming edge is deleted and the counter is decremented, the node is considered garbage if there are no weak incoming edges. If a node contains weak incoming edges after last strong incoming edge deletion,then the algorithm performs a partial tracing to identify if it is a cyclic garbage. Tracing starts by converting all the weak incoming edges of the node into strong by toggling the boolean value. Once the boolean value is changed, the node swaps the counters and identifies all the incoming edges are strong. To identify if the subgraph is cyclic garbage, it weakens all the outgoing edge. The node marks itself that it has performed a transformation. By weakening the outgoing edges mean converting all the outgoing edges of a node into weak edges. Some nodes might lose all strong incoming edges due to this process. This propagates the weakening process. When a node receives a weakening message that already marked the change since the start will decide it is garbage and deletes itself. Every weakening call propagates by switching the weak incoming edges into strong. This method identifies if there are any external references pointing to the node that lost the strong incoming edge.

@@

Figure ~\ref{brownbridge} shows an example graph where Brownbridge technique will identify the external support to a node and restores the graph. In figure ~\ref{brownbridge}, if strong edge to A is deleted, A will convert all incoming weak edges into strong and starts weakening the edges. When B receives weaken message, it weakens the edge and then converts all weak incoming edges to strong and weakens the edge to C. The process continues until it reaches A again. When the process reaches A, A realizes now it has an additional strong incoming edge. So the graph is recovered now. If there is no edge from E to A, then the whole subgraph A, B, C and D will be deleted.

\section{Pitfalls of Brownbridge Garbage Collection}
The idea behind Brownbridge method is to identify external support for the node that lost the last strong incoming edge. If the external strong support to the subgraph comes to a node that is not the node that lost the strong incoming edge, then the method fails to identify them and prematurely delete nodes.

@@

In figure ~\ref{brownbridgefail}, when root deletes edge to A, A starts partial tracing. When the partial tracing reaches back A, D has additional strong support. But algorithm only makes a decision based on A's strong incoming edge. So the A, B, C, D subgraph will be deleted prematurely.

>>>>

The heuristic idea mentioned above prevents the creation of strong cycles. The other technique required by the Brownbridge algorithm is the ability to change the strong and weak edges whenever required by the nodes. To solve this issue, Brownbridge used two boolean values, one for each edge in the source and one for each node. For an edge to be strong, both the boolean values have to be same. So if a node wants to change the type of its edges, a node changes the corresponding boolean value. A node will change all the incoming edge types when it changes its boolean value. By changing the boolean value for an outgoing edge, a node changes the type of the edge. The idea of toggling the edge types is an important piece of the Brownbridge algorithm.

\section{Brownbridge Garbage Collector}
Brownbridge's algorithm works for a system where the Adversary stops the world for the garbage collection process. Each node contains two counters each corresponding to the type of incoming edge. When the graph mutations are serialized, any addition / deletion of edge increments / decrements the corresponding counters in the target node. An addition event never makes a node garbage. A deletion event might make a node garbage. If the edge deleted is weak, no further actions are taken except decrementing the counter. Only a last strong incoming edge deletion to a node performs additional work. When the last strong incoming edge is deleted and the counter is decremented, the node is considered garbage if there are no weak incoming edges. If a node contains weak incoming edges after last strong incoming edge deletion, then the algorithm performs a partial tracing to identify if it is cyclic garbage. Tracing starts by converting all the weak incoming edges of the node into strong by toggling the boolean value. Once the boolean value is changed, the node swaps the counters and identifies all the incoming edges as strong. To determine whether the subgraph is cyclic garbage, the node weakens all its outgoing edges. The node marks itself that it has performed a transformation. By weakening the outgoing edges, we mean converting all the outgoing edges of a node into weak edges. Some nodes might lose all strong incoming edges due to this process. This propagates the weakening process. When a marked node receives a weakening message it will decide it is garbage and deletes itself. Every weakening call propagates by switching the weak incoming edges into strong. This method determines if there are any external references pointing to the node that lost the strong incoming edge.

@@

Figure ~\ref{brownbridge} shows an example graph where the Brownbridge technique will identify the external support to a node and restore the graph. In figure ~\ref{brownbridge}, if the strong edge to A is deleted, A will convert all incoming weak edges into strong and start weakening the outgoing edges. When B receives a weaken message, it weakens the outgoing edge and then converts all weak incoming edges to strong. The process continues until it reaches A again. When the process reaches A, A realizes now it has an additional strong incoming edge. So the graph is recovered now. If there is no edge from E to A, then the whole subgraph A, B, C and D will be deleted.

\section{Pitfalls of Brownbridge Garbage Collection}
The idea behind the Brownbridge method is to identify external support for the node that lost the last strong incoming edge. If the external strong support to the subgraph comes from a node that did not loose its last strong incoming edge, then the method fails to identify the support and prematurely deletes nodes.

@@

In figure ~\ref{brownbridgefail}, when the root deletes the edge to A, A starts partial tracing. When the partial tracing reaches back  to A, D has additional strong support. But the algorithm only makes a decision based on A's strong incoming edge. So the A, B, C, D subgraph will be deleted prematurely.
